\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amscd}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{listings}
\usepackage{enumerate}

\textwidth=15cm \textheight=22cm \topmargin=0.5cm \oddsidemargin=0.5cm \evensidemargin=0.5cm

\newcommand{\sk}{\vskip 10mm}
\newcommand{\bb}[1]{\mathbb{#1}}
\newcommand{\ra}{\rightarrow}

\theoremstyle{plain}
\newtheorem{problem}{Problem}
\newtheorem{lemma}{Lemma}[problem]

\theoremstyle{remark}
\newtheorem{tpart}{}[problem]
\newtheorem*{ppart}{}

\begin{document}

\begin{problem}
  Consider the 3-sphere as the set of unit quaternions
  \[ \{x+iy+jz+kw|x^2+y^2+z^2+w^2=1\} \]
  Let $\phi:U\rightarrow\bb{R}^3$ be $\phi(x+iy+jz+kw)=(y,z,w)\in\bb{R}^3$ where
  \[ U=\{x+iy+jz+kw\in S^3|x>0\} \]
  Consider $\phi$ as a chart. For each $q\in S^3$, define a chart $\psi_q$ by
  $\psi_q(p)=\phi(q^{-1}p)$. Show that this set of charts is an atlas for a
  smooth structure on $S^3$.
\end{problem}

First note that every point $q\in S^3\subset\bb{H}$ will be included
in the domain of its corresponding chart $\psi_q$. Moreover
since $\psi_q$ will be a homeomorphism if $\phi$ is a homeomorphism as $p\mapsto q^{-1}p$
is continuous where defined. Thus all we need to show is that $\phi$ is 
a homeomorphism and that for $q,p\in S^3$ that $\psi_q\psi_p^{-1}$ is smooth.

The projection map $\phi$ is continuous. Define an inverse via
\[ \phi^{-1}(y,z,w)=\sqrt{1-y^2-z^2-w^2}+iy+jz+kw \]
This map is continuous as all of its components are continuous. Thus,
since $\phi$ has a continuous inverse, it is indeed a homeomorphism.
It then follows that $\psi_q$ is a homeomorphism due to the reasoning above.

Now we will show that for $q,p\in S^3$ that $\psi_q\psi_p^{-1}$ is smooth where
defined. Let $r=x+iy+jz+kw$ and $p=a+ib+jc+kd$.
Then we expand $\psi_q\psi_p^{-1}$ as
\begin{align*}
  \psi_q\psi_p^{-1}(r)&=\psi_q(p\phi^{-1}(r))\\
                &= \psi_q(p(\sqrt{1-y^2-z^2-w^2}+iy+jz+kw))\\
%                &= \psi_q((a\sqrt{1-y^2-z^2-w^2}-by-cz-dw)+(ay+\sqrt{1-y^2-z^2-w^2}b+cw-dz)i
%                  +(az-bw+c\sqrt{1-y^2-z^2-w^2}+dy)j+(aw+bz-cy+d\sqrt{1-y^2-z^2-w^2})k)\\
                &= \phi(q^{-1}p(\sqrt{1-y^2-z^2-w^2}+iy+jz+kw))\\
\end{align*}

The expression $q^{-1}p(\sqrt{1-y^2-z^2-w^2}+iy+jz+kw)$ is smooth where defined
and the projection $\phi$ is smooth. Therefore $\psi_q\psi_p^{-1}$ is smooth and
therefore the $\psi_q$s form an atlas for $S^3\subset\bb{H}$.

\sk

\begin{problem}
  Let $X$ be a copy of the real line $\bb{R}$ and let $\phi:X\rightarrow\bb{R}$ be
  $\phi(x)=x^3$. Taking $\phi$ as a chart, this defines a smooth structure on
  $X$. Prove or disprove the following statements:
  \begin{enumerate}
  \item $X$ is diffeomorphic to $\bb{R}$;
  \item the identity map $X\rightarrow\bb{R}$ is a diffeomorphism;
  \item $\phi$ together with the identity map comprise an atlas;
  \item on the one-point compactification $X^+$ of $X$, $\phi,\psi$ give an
    atlas, where $\psi(x)=1/x$, for $x\neq 0,\infty,$ and $\psi(\infty)=0$. ($\psi$ is defined
    on $X^+\setminus\{0\}$.)
  \end{enumerate}
\end{problem}

\begin{enumerate}
\item The smooth manifold $X$ will be diffeomorphic to $\bb{R}$ by the map
  $\phi$. To verify this we check that
  \[ id_{\bb{R}}\circ x^3\circ x^{-1/3}=x\]
  and
  \[ x^3\circ x^{1/3}\circ id_{\bb{R}}^{-1}=x\]
  are smooth. However since they are both the identity map we have that it is
  the case that $\phi$ and $\phi^{-1}$ are both smooth. Thus $\phi$ is a diffeomorphism
  and it follows that $X$ and $\bb{R}$ are diffeomorphic.
  
\item The identity map will not be a diffeomorphism as
  \[ id_{\bb{R}}\circ id\circ x^{1/3}=x^{1/3}\]
  is not a smooth map and thus $id$ cannot be a diffeomorphism.
  
\item Together they do not form an atlas as
  \[ id\circ x^{1/3}=x^{1/3}\]
  is not smooth at zero.
  
\item Yes, the maps $\phi,\psi$ form an atlas on $X^+$. To verify this note
  first that every point in $X^+$ is in the domain of $\phi$ or $\psi$,
  $X$ and $X^+\setminus\{0\}$ respectively. We also have that $x^3$ and $\frac{1}{x}$ are
  homeomorphisms on their respective domains. Finally the maps
  \[ \phi\psi^{-1}=x^{-3}\]
  and
  \[ \psi\phi^{-1}=x^{-1/3}\]
  are both smooth so long as we avoid both $0$ and $\infty$.

  Therefore $\phi$ and $\psi$ form an atlas for $X^+$.
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}
