\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amscd}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{listings}
\usepackage{enumerate}

\textwidth=15cm \textheight=22cm \topmargin=0.5cm \oddsidemargin=0.5cm \evensidemargin=0.5cm

\newcommand{\sk}{\vskip10mm}
\newcommand{\bb}[1]{\mathbb{#1}}
\newcommand{\ra}{\rightarrow}

\theoremstyle{plain}
\newtheorem{problem}{Problem}
\newtheorem{lemma}{Lemma}[problem]

\theoremstyle{remark}
\newtheorem{tpart}{}[problem]
\newtheorem*{ppart}{}

\begin{document}

\begin{problem}
  Consider the real valued function $f(x,y,z)=(2-(x^2+y^2)^{1/2})^2+z^2$
  on $\bb{R}^3\setminus\{(0,0,z)\}$. Show that $1$ is a regular value of $f$.
  Identify the manifold $M=f^{-1}(1)$.
\end{problem}

First we calculate the partials
\[
  \frac{\partial f}{\partial x}=x\left(2-\frac{4}{(x^2+y^2)^{1/2}}\right) ,
  \quad\frac{\partial f}{\partial y}=y\left(2-\frac{4}{(x^2+y^2)^{1/2}}\right) ,
  \quad\frac{\partial f}{\partial z}= 2z 
\]

The critical points of $f$ are those for which the rank of $f_*$ is zero. This
requires that $z=0$. So if $1$ were to be a critical value it would have
to occur when $(x^2+y^2)^{1/2}$ is either $1$ or $3$. However neither of these would
force all the partial derivatives to be zero. As such $1$ is a regular value of $f$.

The manifold is the torus $T^2$.

\sk

\begin{problem}
  Show that the manifold $M$ of Problem 1 is transverse to the surface
  \[
    N=\{(x,y,z)\in\bb{R}^3|x^2+y^2=4\}
  \]
  Identify the manifold $M\cap N$.
\end{problem}

The points of intersection are
\[
  N\cap M = \{(x,y,z)|z=\pm 1, x^2+y^2=4\}
\]

Given any point $p\in\bb{R}^3$ the tangent space $T_p(\bb{R}^3)$ is equivalent
to $\bb{R}^3$. Define the function $h_+:T^2\rightarrow\bb{R}$ as the height of a point
$p\in T^2$ with respect the given embedding. We can write this as
\[
  h_+(x,y)= \sqrt{-{\left(\sqrt{x^{2} + y^{2}} - 2\right)}^{2} + 1}
\]

By taking the partial derivatives with respect to $x$ and $y$ we get
\[
\frac{\partial f}{\partial x} = -\frac{x {\left(\sqrt{x^{2} + y^{2}} - 2\right)}}{\sqrt{x^{2} + y^{2}} \sqrt{-{\left(\sqrt{x^{2} + y^{2}} - 2\right)}^{2} + 1}},\quad \frac{\partial f}{\partial y}=-\frac{y {\left(\sqrt{x^{2} + y^{2}} - 2\right)}}{\sqrt{x^{2} + y^{2}} \sqrt{-{\left(\sqrt{x^{2} + y^{2}} - 2\right)}^{2} + 1}}
\]

which are both zero when $x^2+y^2=4$. Thus the height is not changing on that circle and
it then follows that the tangent plane for such a point is parallel to the $xy$-plane. However
since $N$ is a cylinder in the $z$ direction the tangent plane at any point will include
the $z$ direction. As such at any point of the intersection of $M,N$ for $z>0$ we have
that $TM+TN=\bb{R}^3$. Identical reasoning follows for $z<0$. Thus the manifolds
$N$ and $M$ are transverse.

The manifold is two disjoint copies of $S^1$.
\sk

\begin{problem}
  Show that the manifold $M$ of Problem 1 is not transverse to the surface
  \[
    N=\{(x,y,z)\in\bb{R}^3|x^2+y^2=1\}
  \]
  Is $M\cap N$ a manifold?
\end{problem}

The points of intersection are
\[
  N\cap M = \{(x,y,z)=z=0,x^2+y^2=1\}
\]

Using the partials we calculated in problem one. The normal vector for
any point $(x,y,0)\in T^2$ along the intersection will be $\langle-2x,-2y,0 \rangle$.
Similarly for $N$ the normal vector will be $\langle 2x,2y,0\rangle$. However the
planes defined by these normal vectors will be parallel and as such their
sum cannot be $\bb{R}^3$. Thus the manifolds $M$ and $N$ are not transverse.

It is a manifold and it is a single copy of $S^1$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}
